---
title: "DataExploreNov5"
author: "Ben Szeto"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(gganimate)


```

```{r}

#data_ts<-read.csv("/Users/benszeto/Library/CloudStorage/OneDrive-Personal/Documents/US Elections/anes_timeseries_cdf_csv_20220916/anes_timeseries_cdf_csv_20220916.csv") #windows
data_ts<-read.csv("D:/DocumentsDisk2/US Elections F24/USElectionTEDTalk/anes_timeseries_cdf_csv_20220916/anes_timeseries_cdf_csv_20220916.csv")



```

#General Notes:

VCF0128 RESPONDENT - RELIGION MAJOR GROUP
Question 1952-1964: Is your Church (1962: religious) preference Protestant,
Catholic or Jewish?
1966-1968: Are you Protestant, Catholic or Jewish?
1970-1988,2002: Is your religious preference Protestant, Catholic,
Jewish, or something else?
1990 AND LATER, exc. 2002: (IF R ATTENDS RELIGIOUS SERVICES:) Do you
mostly attend a place of worship that is Protestant, Roman Catholic,
Jewish or what? (IF R DOESN’T ATTEND RELIGIOUS SERVICES:)
Regardless of whether you now attend any religious services do you
ever think of yourself as part of a particular church or
denomination? (IF YES:) Do you consider yourself Protestant, Roman
Catholic, Jewish or what?


Valid 
1. Protestant
2. Catholic [Roman Catholic]
3. Jewish
4. Other and none (also includes DK preference)

```{r}

data_ts<-data_ts%>%
  mutate(Religion_Major_recode=case_when(VCF0128==1~"Protestant",
                                         VCF0128==2~"Catholic(Roman Catholic)",
                                         VCF0128==3~"Jewish",
                                         VCF0128==4~"Other and None",
                                         VCF0128==0~"Missing")
  )
```

```{r}

Religion_Major_Year<-data_ts%>%
  group_by(VCF0004,Religion_Major_recode)%>%
  summarize(
    n()
  )%>%
  group_by(VCF0004)%>%
  mutate(percent=`n()`*100/sum(`n()`))

#Religion_Major_Year

Religion_Major_Year%>%
  ggplot(aes(x=VCF0004, y=percent, color=Religion_Major_recode))+
  geom_line()+
  geom_point()+
  labs(x="Year", y="Percent", title="Major Religions over Time")#+
  #transition_reveal(VCF0004) #make it animated
```





VCF0130 RESPONDENT - CHURCH ATTENDANCE 1970-

1. Every week (Except 1970: almost every week)
2. Almost every week (no cases in 1970)
3. Once or twice a month
4. A few times a year
5. Never (1990 and later: ‘No’ in filter)
7. No religious preference (1970-1988)
Missing 8. DK how often/DK if attend
9. NA/RF how often/NA/RF if attend
0. Religion (1970-1988); atheists and agnostics are Inap (1972,1986 only);
short-form ‘new’ Cross Section (1992-1996)
INAP. question not used

```{r}

data_ts<-data_ts%>%
  mutate(Church_Attendence_recode=case_when(VCF0130%in%c(1)~"Every Week",
                                            VCF0130%in%c(2,3)~"At least once a month",
                                            VCF0130%in%c(4)~"A few times a year",
                                            VCF0130%in%c(5,7)~"Never"
                                            
         ))



Church_Attendence_Year<-data_ts%>%
  filter(VCF0004>1970)%>%
  filter(Church_Attendence_recode%in%c("Every Week", "At least once a month", "A few times a year", "Never"))%>%
  group_by(VCF0004, Church_Attendence_recode)%>%
  summarize(
    n()
  )%>%
  group_by(VCF0004)%>%
  mutate(percent=`n()`*100/sum(`n()`))

Church_Attendence_Year

Church_Attendence_Year%>%
  ggplot(aes(x=VCF0004, y=percent, color=Church_Attendence_recode))+
  geom_line()+
  geom_point()+
  labs(x="Year", y="Percent", title="Church Attendence")+
  transition_reveal(VCF0004)

```


```{r}
data_ts<-data_ts%>%
  mutate(Church_Attendence_recode=case_when(VCF0128==1~"Every Week",
                                         VCF0128==2~"Almost Every Week",
                                         VCF0128==3~"Once or Twice a Month",
                                         VCF0128==4~"A few times a year",
                                         VCF0128==5~"Never",
                                         VCF0128==7~"No religious Preference",
                                         )
  )

```


```{r}

Church_Attendence_Year<-data_ts%>%
  group_by(VCF0004,Religion_Major_recode)%>%
  summarize(
    n()
  )%>%
  group_by(VCF0004)%>%
  mutate(percent=`n()`*100/sum(`n()`))

#Religion_Major_Year

Religion_Major_Year%>%
  ggplot(aes(x=VCF0004, y=percent, color=Religion_Major_recode))+
  geom_line()+
  geom_point()+
  labs(x="Year", y="Percent", title="Major Religions over Time")+
  transition_reveal(VCF0004) #make it animated
```




VCF0302 PARTY IDENTIFICATION OF RESPONDENT- INITIAL PARTY
ID RESPONSE



```{r}



```


What religions go with each party (1980 vs 2024)

evangelicals


```{r}

Protestant_Evangelical_1984_PartyIdentification<-data_ts%>%
  select(VCF0004, VCF0128a, VCF0302)%>%
  mutate(Party_Identification_recode=case_when(VCF0302==1~"Republican",
                                               VCF0302%in%c(2,3,4)~"Independent,Other,  No preference/none/neither",
                                               VCF0302==5~"Democrat"))%>%
  filter(VCF0004==1984)%>%
  filter(VCF0128a==2)%>%#
  filter(!is.na(Party_Identification_recode))%>%
  count(Party_Identification_recode)
  
 Protestant_Evangelical_1984_PartyIdentification%>%
  ggplot(aes(x="", y=n, fill=Party_Identification_recode))+
  geom_bar(stat="identity", width=1)+
  coord_polar("y", start=0)+
  theme_void()
   

```




2020 Dataset


```{r}
data_2020<-read.csv("D:/DocumentsDisk2/US Elections F24/USElectionTEDTalk/anes_timeseries_2020_csv_20220210/anes_timeseries_2020_csv_20220210.csv")

```


V201206 - Where would you put the democratic party of a scale
V201230- if no party closer republican or democrat

```{r}
data_2020%>%
  select(V201458x)%>%
  count(V201458x)


Protestant_Evangelical_2020_PartyReg<-data_2020%>%
  mutate(Party_Registration_recode=case_when(V202065x==1~"Democratic Party",
                                              V202065x==2~"Republican Party",
                                              V202065x%in%c(4,5)~"Other Party or Independent",
                                              V202065x<0~"Missing"))%>%
  filter(!Party_Registration_recode=="Missing")%>%#Removing missing
  filter(V201458x==2)%>%#selecting evangelical protestants
  count(Party_Registration_recode)

Protestant_Evangelical_2020_PartyReg

Protestant_Evangelical_2020_PartyReg%>%
  ggplot(aes(x="", y=n, fill=Party_Registration_recode))+
  geom_bar(stat="identity", width=1)+
  coord_polar("y", start=0)+
  theme_void()
  

```

2012 data
```{r}
library(haven)
data_2012<-read_dta("D:/DocumentsDisk2/US Elections F24/USElectionTEDTalk/anes_timeseries_2012_dta/anes_timeseries_2012.dta")

```

```{r}
Protestant_Evangelical_2012_PartyReg<-data_2012%>%
  mutate(Party_Registration_recode=case_when(prevote_regpty==1~"Democratic Party",
                                              prevote_regpty==2~"Republican Party",
                                              prevote_regpty%in%c(4,5)~"Other Party or Independent",
                                              prevote_regpty<0~"Missing"))%>%
  filter(relig_7cat_x==2)%>%#Protestant Evangelicals
  filter(!Party_Registration_recode=="Missing")%>%#Removing Missing
  count(Party_Registration_recode)#Party of registration #What political party are you registered with, if any?



Protestant_Evangelical_2012_PartyReg%>%
  ggplot(aes(x="", y=n, fill=Party_Registration_recode))+
  geom_bar(stat="identity", width=1)+
  coord_polar("y", start=0)+
  theme_void()
  

```



1984 Data (this data seems to be pretty limited)


```{r}

data_1984<-read_dta("D:/DocumentsDisk2/US Elections F24/USElectionTEDTalk/anes1984dta/NES1984.dta")

data_1984

```

```{r}

data_1984%>%
  count(V840687)#only 7 evangelical protestants?

```